Big Data with High Performance Computing (HPC)

### 2016-04-03 Tex_Files/paper.tex
HPC MapReduce added. Graphic recreated.

\usepackage{graphicx}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\blindtext
With the recent emergence of Big Data over the past few years, one of the questions that arises is, how Big Data is different from High Performance Computing (HPC) at various levels such as application, technology, investment, etc. Another critical question to be asked is if Big Data is widely different from HPC, which of the technologies may we want to adopt or, do we really have to choose one from the two to apply to our problems? Another scenario is that the gap between Big Data and HPC may form a spectrum of choices rather than two independent worlds. To answer these questions in this chapter we will first bring an introduction and look at Big Data and HPC from a high level abstraction view. Then, we explain the state of the art efforts to take advantage of the best of two worlds. Finally, we will address future challenges and future directions.\\

\begin{figure}
\includegraphics[scale=0.31]{./images/BigData_HPC.png}
\centering
\caption{Matrix of data and computation applications.}
\label{fig:bigdata_hpc_matrix}
\end{figure}
In order to answer the above questions we first may need to identify the workloads and computational requirements of workflows for our applications. It is also important to not focus only current trends, and take into account the future data and computation demands. Figure \ref{fig:bigdata_hpc_matrix} shows the matrix of data size versus computation size for some applications. It also shows that the current big data technologies, formed around Apache Big Data Stack (ABDS), is data-intensive but focused less on computational performance. On the other hand the High Performance Computing (HPC), which mostly applied to solving scientific problems, has been mostly concentrated on computation-savvy, data limited application such and iterative simulations.\\

Over the past decade, the rate of data generation has dramatically increased. As a result, we saw the emergence of data-intensive applications to solve a range of problems from data  distribution and management to data processing. The open-source Apache Big Data Stack with it's Hadoop Data File System (HDFS) kernel and many higher abstraction level frameworks for data analytics and machine learning has provided a coherent ecosystem, which has dominated the industry.\\

In contrast, HPC with roots in computational intensive scientific computing, has mainly been applied and fine-tuned for particular problems, which resulted in significantly more limited implementations compared to ABDS.
\begin{figure}
	\includegraphics[scale=0.33]{./images/Ecosystems.png}
	\centering
	\caption{Apache Big Data Stack (ABDS) ecosystem vs. High Performance Computing (HPC) ecosystem \cite{Jha2014}.}
	\label{fig:bigdata_hpc_ecosystem}
\end{figure}
Figure \ref*{fig:bigdata_hpc_ecosystem} depicts the HPC and Big Data ecosystems from many abstraction level view. The lowest layers are the resources, resource management, and communication layers. The higher layers comprise of processing and analytical components. In remaining parts of this section we compare current typical HPC and Big Data architectures and their respective ecosystems.
\subsection{HPC Architecture}
A typical HPC architecture consists of processing units such a multi-core structures, possibly accelerated by GPUs, FPGAs, or even CPUs such as intel Xeon Phi. These processing units are supported by local storage units with Lustre \cite{braam2004lustre} or General Parallel File System (GPFS) \cite{FrankSchmuck} file system or a remote storage connected through a fast network such as Infiniband.\\

@@

In HPC applications, data need to fly around across network using a low latency communication. These fast interconnect networks along with features such as non-blocking and one-sided communications allow for more data-intensive HPC within HPC environment. As a matter of fact, data-intensive applications are not totally unfamiliar to HPC community. There has been many approaches to implement MapReduce compatible with HPC environment such as MapReduce-MPI \cite{plimpton2011mapreduce}, Pilot-MapReduce \cite{mantha2012pilot}, and Twister for machine learning applications \cite{ekanayake2010twister}.In a parallel effort, there has been many implementations of data intensive loosely coupled tasks at run-time level \cite{luckow2012p}\cite{raicu2008many}\cite{deelman2009workflows}.
\subsection{ABDS Architectures}
The Apache Big Data Stack is formed around the Hadoop Distributed File System (HDFS), which originated at Yahoo but is an open source implementation of Google's file system \cite{shvachko2010hadoop} \cite{ghemawat2003google}. As we know now, the idea behind ABDS was to bring the computation to data, which is usually stored at cheap commodity hardware. Hadoop 1.0 also took advantage of Google's MapReduce programming model for data processing. MapReduce imposes it's own limitations since all processes is forced to be implemented in the form of a mapper followed by a reducer. MapReduce's inherent limitations along with tight coupling of Hadoop and MapReduce proved to be inflexible when used in applications such as iterative computations, which is an indispensable piece of all Machine Learning algorithms.\\

As a result of such deficits in Hadoop 1.0, Apache YARN was introduced with Hadoop 2.0 as a much more lenient resource manager, enabling many applications and frameworks with higher level abstractions \cite{vavilapalli2013apache}. YARN's key feature was introduction multi-level scheduling, which would allow higher level applications to schedule their own processes. As a results we saw the emerge of many high level applications such as, HBase, a column based distributed database, Spark, Giraph, which are iterative and graph processing applications\cite{borthakur2011apache}\cite{zaharia2012resilient}.
\newpage
\section{State of the art projects}
In this section we will review some of the state of the art researches and projects within HPC-Big Data domains. We categorized these project under four class of research topics, namely,
\begin{enumerate}
	\item Integration of HPC and ABDS stack
	\item Machine Learning (ML) using HPC
	\item Parallel databases
	\item Parallel I/O

\end{enumerate}
\subsection{Integration of HPC and Big Data stack}
Big Data and HPC used to be used interchangeably for many years. But, with HPC targeting scientific problems with huge number of iterations, and Big Data aiming for relatively simple model on huge datasets, it seems they diverged for a while. As the rate of data generation increase, deployment of HPC tools and techniques seem more and more inevitable. Following projects are some of the state of the arts in combining HPC and Big Data stack.
\subsubsection{High Performance Big Data System (HPBDS) project \cite{qiu2014towards}}: The HPBDS project aims for integrating some of the components of the HPC such as scientific libraries, fast communication and resource management features with commercial Big Data ecosystem namely, ABDS.\\

The primary proposed execution of HPBDS is built around Apache Hadoop and therefore named High Performance Computing Big Data Stack (HPC-ABDS). The HPC-ABDS comprise two sub-project,
\begin{figure}[h]
	\includegraphics[scale=0.25]{./images/MIDASSPIDAL.png}
	\centering
	\caption{MIDAS and SPIDAL within HPC-ABDS stack\cite{qiu2014towards}}
	\label{fig:MIDAS_SPIDAL}
\end{figure}
\begin{enumerate}
	\item Middleware for Data Intensive Analytics and Science (MIDAS)
	\item Scalable Parallel Interoperable Data Analytics Library (SPIDAL)  
\end{enumerate}

Figure \ref{fig:MIDAS_SPIDAL} depicts the structure of the MIDAS-SPIDAL with trespect to the overall HPC-ABDS stack. MIDAS is proposed to provide a lower level infrastructure based on which higher level components such as libraries operates. On the other hand SPIDAL goal is to take advantage of useful tool within HPC such as MPI, PETSc, and SCALAPACK and modify them for data intensive applications.\\

The HPC-ABDS project is inspired by the NIST Big Data initiative in Fall 2013, when NIST provided a collection of use cases. These use cases were analyzed against the identifiers proposed as big data Ogres \cite{fox2014towards}. The result is tabulated in Table \ref{table:NIST_vs_ogres}.
\begin{table}
	\centering
	\caption{Categories of the NIST use cases.}
	\begin{tabular}{ |p{2cm}|p{1.5cm}|p{8cm}|  }
		\hline
		Abbrevation & Count & Description \\
		\hline
		PP & 26 & Pleasingly Parallel or Map Only \\
		\hline
		MR & 18 & Classic MapReduce MR (add MRStat below for full count) \\
		\hline
		MRStats & 7 & Simple version of MR where key computations are simple reduction as found in
statistical averages such as histograms and averages \\
\hline
MRIter & 23 & Iterative MapReduce or MPI \\
\hline
Graph & 9 & Complex graph data structure needed in analysis \\
\hline
Fusion & 11 & Integrate diverse data to aid discovery/decision making; could involve sophisticated
algorithms or could just be a portal \\
\hline
Streaming & 41 & Some data comes in incrementally and is processed this way \\
\hline
Classify & 30 & Classification: divide data into categories\\
\hline
S/Q & 12 & Index, Search and Query\\
\hline
CF & 4 & Collaborative Filtering for recommender engines \\
\hline
LML & 36 & Local Machine Learning (Independent for each parallel entity) \\
\hline
GML & 26 & Global Machine Learning: Deep Learning, Clustering, LDA, PLSI, MDS, Large
Scale Optimizations as in Variational Bayes, MCMC, Lifted Belief Propagation,
Stochastic Gradient Descent, L-BFGS, Levenberg-Marquardt. Can call EGO or
Exascale Global Optimization with scalable parallel algorithm \\
\hline
 & 51 & Workflow: Universal so no label \\
\hline 
 GIS & 16 & Geotagged data and often displayed in ESRI, Microsoft Virtual Earth, Google
Earth, GeoServer etc. \\
\hline
HPC & 5 & Classic large-scale simulation of cosmos, materials, etc. generating (visualization)
data \\
\hline
Agent & 2 & Simulations of models of data-defined macroscopic entities represented as agents \\
\hline

	\end{tabular}
	\label{table:NIST_vs_ogres}
\end{table}
Based on the the tabulated results od the table \ref{table:NIST_vs_ogres}, Qiu et al. \cite{Qiu2014} identified five programming models to be used programs addressed in big data HPC problems. These five programming models are shown in table \ref{table:NIST_5_Programming_model}.
%%%%%%%%%%%%%%%%%%%% EXACTT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
	\centering
	\caption{Five programming models identified for big data HPC problems based on NIST big data use cases.}
	\begin{tabular}{ |p{4cm}|p{7cm}|  }
		\hline
		\textbf{Programming Model} & \textbf{Description} \\
		\hline
		Pleasingly Parallel & Applications such as local machine learning where processing is applied over many items in a local machine. Either Hadoop or other HPC tools could be used. \\
		\hline
		Classic MapReduce & Including MRStat, some search applications, collaborative filtering and motif finding. Traditional Hadoop and MapReduce could be used.\\
		\hline
		Iterative Map Collective & Iterative applications of MapReduce using collective communication. Hadoop 2.0 along with tools such as Spark could be used.\\
		\hline
		Iterative Map Communication & Iterative MapReduce such as Giraph with pointto-point
		communication, includes most graph algorithms
		such as maximum clique, connected
		component, finding diameter, community detection).
		 \\
		\hline
		Shared Large Memory & Thread-based (event driven) graph algorithms
		such as shortest path and Betweenness centrality.
		Large memory applications.\\
		\hline

@@

\newpage
\subsection{Machine Learning and HPC}
Traditionally the words Machine Learning and HPC would imply two separate worlds with distinct border. But, over the past decade areas of AI such as machine learning and training neural networks of deep learning have found their ways through HPC world.\\

Iterative AI tools such as machine learning and deep learning initially introduced to HPC by adopting GPU accelerators and FPGAs. Such tools adopted to perform a wide range of AI task from image classification, audio recognition, data analytics. Machine learning is not a novel concept and has been around for a relatively long time. But, along with more computational power introduced by accelerators, two major factors played an undeniable role in re-emergence of machine learning and deep learning. First, data generation at the very fast rates fuels up machine learning training algorithms. In a series of reports by Domo, \cite{domo} the amount of data created at the some of the biggest Internet giants, is presented. Table show some of these numbers in only one minute time of Internet. Secondly, introduction of new training algorithms such as those adopted to train deep belief nets and resulted to the birth of the term Deep Learning \cite{hinton2006fast}. Only even twenty years ago it would take very long time, if possible, to train a neural network with only ten hidden layers.
\begin{table}
	\centering
	\caption{One minute of Internet usage \cite{domo}.}
	\begin{tabular}{ |p{3cm}|p{9cm}|  }
		\hline
		Company & Count \\
		\hline
		Facebook & Users like \textbf{4,166,667} posts.\\
		\hline
		Twitter & Users send \textbf{347,222} tweets.\\
		\hline
		Youtube & Users upload \textbf{300} hours of video.\\
		\hline
		Instagram & Users like \textbf{1,736,111} photos. \\
		\hline
		Pinterest & Users pin \textbf{9,722} images.\\
		\hline
		Apple & Users download  \textbf{51,000} apps.\\
		\hline
		Netflix & Subscribers stream  \textbf{77.160} hours of video.\\
		\hline
		Reddit & Users cast  \textbf{18,327} votes.\\
		\hline
		Amazon & Receives  \textbf{4,310} unique visitors.\\
		\hline
		Vine & Users play  \textbf{1,041,666} videos.\\
		\hline
		Tinder & Users swipe  \textbf{590,276} times.\\
		\hline
		Snapchat & Users share  \textbf{284,722} snaps.\\
		\hline
		Buzzfeed & Users view  \textbf{34,150} videos.\\
		\hline
		Skype & Users make  \textbf{110,040} calls.\\
		\hline
		Uber & Passengers take  \textbf{694} rides.\\
		\hline

@@

Deep Learning, which itself is a sub-category of machine learning. The major distinguishing difference between most of machine learning techniques and deep learning is in feature selection stage. In conventional machine learning techniques labeled/unlabeled data is fed into a set of hand crafted feature detectors to extract features. A classifier takes the features as inputs to perform some statistical analysis and decide on the class of each input. Since the engineering of a good set of such hand crafted features is extremely costly and require a lot of expertise and domain knowledge, deep learning is introduced to find and tune such features automatically using machine and data to be used for a particular task.\\

There is a general consensus that scaling up each of the the training data size and model size will increase the accuracy of prediction in deep learning. Therefore, as the size of training data or the model increases, we require to parallelize the process on many machines and processing cores to enable the scaling. Consequently, two parallelization schemes is offered namely, data parallelization and model parallelization.\\

>>>>

\usepackage{graphicx}
\usepackage{lipsum}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\blindtext
For a long time the terms Big Data and High Performance Computing (HPC) have been used interchangeably implying that they convey the same meaning. Many experts take this even further and recognize HPC as the entity that gave birth to Big Data. However, the two seem to have chosen separate paths fueled by two different driving forces namely, scientific community and industry. HPC has its roots in computation-intensive scientific modeling performing iterative algorithms. In contrast, Big Data was born as a result of a demand for gaining insight from large unstructured data stored by industries and nowadays by us human beings digital life. But, as data generation rate out-paces its processing workforce, Big Data solutions face an inevitable require to adopt more advance processing approaches having higher efficiency and performance.  As a result the boundary between HPC and Big Data is fading everyday. Authors believe that the gap between Big Data and HPC may form a spectrum of choices rather than two independent worlds. In this chapter we will first bring a brief introduction on Big Data and HPC then we will take and look at big data and HPC from a high level abstraction view. The majority of this chapter is dedicated to state of the art efforts to take advantage of the best of two worlds. Finally, we will address future challenges and future directions.\\

\begin{figure}
\includegraphics[width=\textwidth]{./images/bigdata_matrix.png}
\centering
\caption{Matrix of data and computation applications.}
\label{fig:bigdata_hpc_matrix}
\end{figure}
In order to address the above issues we first may need to identify the workloads and computational requirements of workflows for each application. It is also important to not focus only on current trends, and take into account the future data and computation demands. Figure \ref{fig:bigdata_hpc_matrix} shows the matrix of data size versus computation size for some applications. It also shows that the current big data technologies is data-intensive but focused less on computational performance. On the other hand the High Performance Computing (HPC), which mostly applied to solving scientific problems and fine-tuned for particular problems, has been mostly concentrated on computation-savvy, data limited application such and iterative simulations commonly used to solve scientific problems.\\

\subsection{Big Data Ecosystem}
Big data is commonly distinguished by its 5V properties namely, Volume, Variety, Velocity, Value, and Veracity. However, Vs are not the only factors impacting the architecture of a big data ecosystem. In \cite{demchenko2014defining} Demchenko \textit{et al.} propose a definition of big data characteristics and requirements in five categories. These categories are, 5V properties, new data models, new data analytics, new infrastructures, and new data sources and targets. Therefore, a big data ecosystem should be able to address the problems introduced by each of these properties.\\

It should be noted that there has not been a well accepted definition of big data either by industry or academia. This ambiguity makes it even harder to characterize big data and define requirements of a big data ecosystem. In addition, big data ecosystem is not merely made out of a single or couple of components such as a file system or a database. A big data ecosystem should account for lots of components from storage and processing to orchestration and high level applications. Figure \ref{fig:bigdata_ecosystem} depicts a generic architecture of a big data ecosystem.

\begin{figure}
	\includegraphics[scale=0.23]{./images/bigdata_ecosystem.png}
	\centering
	\caption{A general big data ecosystem architecture.}
	\label{fig:bigdata_ecosystem}
\end{figure}

\subsubsection*{ABDS Architectures}
The Apache Big Data Stack is formed around the Hadoop Distributed File System (HDFS), which originated at Yahoo but is an open source implementation of Google's file system \cite{shvachko2010hadoop} \cite{ghemawat2003google}. As we know now, the idea behind ABDS was to bring the computation to data, which is usually stored at cheap commodity heterogeneous hardware. Apache Hadoop 1.0 adopted Google's MapReduce programming model for data processing. MapReduce imposes its own limitations since all processes are forced to be implemented in the form of a mapper and a reducer. MapReduce's inherent limitations along with tight coupling of Hadoop and MapReduce proved to be inflexible when used in applications such as iterative computations, which is an essential piece of most of higher level analytics such as Machine Learning algorithms. Figure \ref{fig:ABDS_ecosystem} depicts the ABDS ecosystem.\\

\begin{figure}
	\includegraphics[scale=0.23]{./images/abds_ecosystem.png}
	\centering
	\caption{Apache Big Data Stack (ABDS) ecosystem.}
	\label{fig:ABDS_ecosystem}
\end{figure}

As a result of such shortcomings of Hadoop 1.0, Apache YARN was introduced with Hadoop 2.0 as a much more lenient resource manager, enabling many applications and frameworks with higher level abstractions \cite{vavilapalli2013apache}. YARN's key feature was introduction multi-level scheduling, which would allow higher level applications to schedule their own processes. As a results we saw the emerge of many high level applications such as, HBase, a column based distributed database, Spark, Giraph, which are iterative and graph processing applications respectively\cite{borthakur2011apache}\cite{zaharia2012resilient}.

\subsection{HPC Ecosystem}

A typical HPC architecture consists of processing units such a multi-core structures, possibly accelerated by GPUs, FPGAs, or even CPUs such as intel Xeon Phi. These processing units are supported by local storage units with Lustre \cite{braam2004lustre} or General Parallel File System (GPFS) \cite{FrankSchmuck} file system or a remote storage connected through a fast network such as Infiniband. Figure \ref{fig:HPC_ecosystem} shows an example of a typical HPC ecosystem.\\

\begin{figure}
	\includegraphics[scale=0.23]{./images/hpc_ecosystem.png}
	\centering
	\caption{An example of HPC ecosystem.}
	\label{fig:HPC_ecosystem}
\end{figure}

@@

In HPC applications, data need to fly around across network using a low latency communication. These fast interconnect networks along with features such as non-blocking and one-sided communications allow for more data-intensive HPC within HPC environment. As a matter of fact, data-intensive applications are not totally unfamiliar to HPC community. There has been many approaches to implement MapReduce compatible with HPC environment such as MapReduce-MPI \cite{plimpton2011mapreduce}, Pilot-MapReduce \cite{mantha2012pilot}, and Twister for machine learning applications \cite{ekanayake2010twister}.In a parallel effort, there has been many implementations of data intensive loosely coupled tasks at run-time level \cite{luckow2012p}\cite{raicu2008many}\cite{deelman2009workflows}.

\subsection{Big Data Applications}
In the beginning of the twenty first century data mining and data processing became the driving force behind many fields such as business, Information Technology (IT), etc. Over the recent past Big Data has risen to address the many large-scale data driven problems, which formerly viewed as unsolvable. In this part we will briefly discuss these applications in two major environments namely, commercial and scientific environments.\\

\textit{Big Data Commercial Applications:} Since 90s until the recent present, most of the data stored by companies were in the form of relational databases. However, nowadays with the enormous data generated in various forms such as activity logs, pictures, audio, clicks, and lot of other types, old database paradigms proved to be inefficient. As a results big data analytics with non-relational databases, distributed storages, secure and fault tolerant architectures have gained momentum in data analytics from data storing, data cleansing, to complicated statistical analysis for predictive analytics and visualization. In parallel, fields such as Machine Learning and Deep Learning where data is the driving thrust, benefited the most from Big Data. This trend introduced and revived many applications within Artificial Intelligent (AI) realm. The synergy between the AI and dominance of web based applications working with unstructured data has already discovered new universe of applications such as videos surveillance for law enforcements and intelligence agencies, recommendation engines for personalized shopping, etc. \\

\textit{Big Data Scientific Applications:} Scientific communities acquaintance with big data predates industry. Particle physics, Astrophysics, Genetics, Meteorology, and lot of other fields have been dealing with massive amounts of data, most of the times beyond their processing power. Recently, the National Science Foundation (NSF) announced the Big Data research initiative and called the researches to extract knowledge and insight from data.
\subsection{Integration of big data and HPC ecosystems}
Traditional data analytics in the past have been able to extract general high level information from data. But, as the data started to generated with the 5V properties, those old method proved to be less and less effective. Consequently, many data consumers have pointed their short-term and long-term agendas towards implementation and adoption of new methods of data analysis.\\

Despite Big Data and HPC's different contexts, HPC has a relatively long history addressing many of these challenges. HPC is conventionally built around the expensive supercomputers including a large number of servers. On the other hand Big Data has its roots in commodity hardware focusing on priorities such a fault tolerance, scheduling, and heterogeneous computing. HPC components are usually tightly coupled as opposed to its loosely coupled big data counterpart. It is evident that in order to take advantage of the best of the two worlds, we need some breakthroughs in HPC to provide more heterogeneous computing power and Big Data to embrace more efficient components from HPC.\\

In addition, HPC and big data communities have favored different programming models. Due to shortcomings of MPI alone, which is incapable of scaling at Exascale, current trend in HPC is to take advantage of MPI+X as runtime communication libraries. These libraries require to support many MPI+X combinations such as MPI+OpenMP, MPI+UPC, MPI+SHMEM, etc. It should be noted that at Exascale such libraries should be able to provide inter/intra node communications for billion processors.

\section{State of the art projects}
In this section we will review some of the state of the art researches and projects within HPC-Big Data domains. We categorized these project under four class of research topics, namely,
\begin{enumerate}
	\item Integration of HPC and ABDS stack
	\item Machine Learning (ML) using HPC
	\item High Performance Graph processing
	\item Parallel databases
	\item HPC implementation of MapReduce

\end{enumerate}
\subsection{Integration of HPC and big data stack}
Earlier in this chapter we discussed how HPC can offer many solutions to Big Data's current challenges. In this section we review a state of the art effort to combine HPC and Big Data ecosystem under {High Performance big data System (HPBDS) project.

\subsubsection{High Performance Big Data System (HPBDS) project \cite{qiu2014towards}}: The HPBDS project aims for combining some of the components of the HPC such as scientific libraries, fast communication and resource management features with commercial Big Data ecosystem namely, ABDS.\\

The proposed design of HPBDS is built around the commercial Apache Hadoop project and therefore named High Performance Computing Big Data Stack (HPC-ABDS). The HPC-ABDS comprise two sub-project,
\begin{figure}[h]
	\includegraphics[scale=0.25]{./images/hpc_abds.png}
	\centering
	\caption{MIDAS and SPIDAL within HPC-ABDS stack\cite{qiu2014towards}}
	\label{fig:MIDAS_SPIDAL}
\end{figure}
\begin{enumerate}
	\item Scalable Parallel Interoperable Data Analytics Library (SPIDAL)
	\item Middleware for Data Intensive Analytics and Science (MIDAS)  
\end{enumerate}

Figure \ref{fig:MIDAS_SPIDAL} depicts the structure of the MIDAS-SPIDAL with respect to the overall HPC-ABDS stack. MIDAS is proposed to provide a lower level infrastructure based on which higher level components such as libraries operates. MIDAS aims to couple performance of HPC with functionality of heterogeneity of Apache Hadoop tools. On the other hand SPIDAL goal is to take advantage of useful tool within HPC such as MPI, PETSc, and SCALAPACK and modify them for data intensive applications.\\

The HPC-ABDS project is inspired by the NIST big data initiative in Fall 2013, when NIST provided a collection of use cases. These use cases were analyzed against the identifiers proposed as big data Ogres \cite{fox2014towards}. Based on the results from analyzing 71 use cases provided by NIST against data Ogres, five programing models was established. Table \ref{table:NIST_5_Programming_model} summarizes these programming models.

%%%%%%%%%%%%%%%%%%%% EXACTT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
	\centering
	\caption{Five programming models identified for big data HPC problems based on NIST big data use cases.}
	\begin{tabular}{ |p{4cm}|p{7cm}|  }
		\hline
		\textbf{Programming Model} & \textbf{Description} \\
		\hline
		Pleasingly Parallel & Applications such as local machine learning where processing is applied over many items in a local machine. Either Hadoop or other HPC tools could be used. \\
		\hline
		Classic MapReduce & Including MRStat, some search applications, collaborative filtering and motif finding. Traditional Hadoop and MapReduce could be used.\\
		\hline
		Iterative Map Collective & Iterative applications of MapReduce using collective communication. Hadoop 2.0 along with tools such as Spark could be used.\\
		\hline
		Iterative Map Communication & Iterative MapReduce such as Giraph with pointto-point
		communication, includes most graph algorithms
		such as maximum clique, connected
		component, finding diameter, community detection).
		 \\
		\hline
		Shared Large Memory & Thread-based (event driven) graph algorithms
		such as shortest path and Betweenness centrality.
		Large memory applications.\\
		\hline

@@

\subsection*{Scalable Parallel Interoperable Data-Analytics Library (SPIDAL) }
SPIDAL project is formed around characteristics identified by analyzing NIST uses cases and programming models in Table \ref{table:NIST_5_Programming_model} to build higher level abstractions based on the middleware analytic tools proposed by MIDAS. The followings comprise  core library category of SPIDAL.
\begin{itemize}
	\item \textbf{Graph and Networks:} With graphs taking over many applications that traditionally did not belong to them, graph processing has become a more and more important. SPIDAL approach for graph processing involve creation of new tools and techniques based on both MapReduce and Giraph \cite{avery2011giraph} model. Graph processing solutions are based on structures withing the graph or network. For a wide range of structures Giraph based solutions seem to better perform mostly when a distributed memory parallel approach is taken. On the other hand a range of graphs respond better to MapReduce solutions when problem can be disintegrated into reduction operations.\\
	\item \textbf{Spacial Queries and Special Analytics:} Spacial Data Management Systems (SDBMS) are incapable of performing tasks that require lots of computations optimally. Thus, the proposed design shards the space into partitions and breaks the most time-consuming partitions into multiple chunks feeding them into parallel MapReduce. Hadoop-GIS \cite{aji2013hadoop} embedded into Apache Hive enables querying on MapReduce.\\

	\item \textbf{Image Processing:} Computer vision over the past few years has been empowered by introduction of new tools and techniques such as GPUs and Learning algorithms. In addition, image processing tasks generally tend to be highly parallelizeable. SPIDAL will provide a comprehensive image processing library both at high system level and primitive pre-processing functions and kernels such as Convolution.\\

	\item \textbf{Machine learning:} There are many Machine Learning libraries designed for Apache Hadoop framework. Within Hadoop ecosystem, Apache Mahout supports many of both supervised and unsupervised Machine Learning algorithms. SPIDAL will include a set of key Machine Learning algorithms such as Support Vector Machines (SVM), Random Forests, etc. SPIDAL will try to have Machine Libraries parallelized. Besides, SPIDAL will give those iterative algorithms that require disk I/O in current Big Data environment, a bigger boost in performance.  
\end{itemize}

\subsection*{Middleware for Data Intensive Analytics and Science (MIDAS)}

MIDAS is proposed to facilitate a lower abstraction level, which includes run time and resource management for SPIDAL. We now know from analysis of big data ogres and NIST use cases that big data application vary widely in requirements and characteristics. Therefore, different programming models and parallel approaches should be considered dealing with each type of problems. With this in mind, MIDAS has targeted two major objectives. First, a high abstraction level environment that supports querying and analysis via HPC and big data tools such as SLURM and YARN as schedulers. Secondly, to provide a middleware that support the four of the identified programming models in table \ref{table:NIST_5_Programming_model} namely, Pleasingly Parallel, Classic MapReduce, Iterative Map with Collectives, and Iterative Map Communication. Hence, difference abstraction layers have been recommended. For instance, in applications that better fit the iterative MapReduce, most operations are collective and MPI would perform better compared to MapReduce techniques. In contrast, graph processing application require more point to point communications.\\

In conclusion, HPC-ABDS as claimed by the authors is the first attempt to create a HPC empowered big data stack. The HPC-ABDS tries to find and characterize application patterns through a set of features named as Ogres. A set of libraries (SPIDAL) are created based on characteristics identified by the Ogres run across various computing platforms. SPIDAL runs on MIDAS, which is a middleware with resource management and data abstraction capabilities.

\subsection{Machine Learning and HPC}
Traditionally the words Machine Learning and HPC would imply two separate worlds with distinct border. But, over the recent past, areas of AI such as machine learning and training neural networks of deep learning have found their ways through HPC world.\\

Iterative AI tools such as machine learning and deep learning initially introduced to HPC through GPU accelerators and FPGAs. These tools were adopted to perform a wide range of AI tasks from image classification, audio recognition, to data analytics. Machine learning is not a novel concept and has been around for a relatively long time. But, along with more computational power introduced by accelerators, two major factors played an undeniable role in re-emergence of machine learning and deep learning. First, data generation at blasting rates fuels up machine learning training algorithms. In a series of reports by Domo, \cite{domo} the amount of data created at the some of the biggest Internet giants, is presented. Table \ref{table:Domo} shows some of these numbers in only one minute time of Internet. Secondly, introduction of new training algorithms, which also gave birth to the term Deep Learning \cite{hinton2006fast}.
\begin{table}
	\centering
	\caption{One minute of Internet usage \cite{domo}.}
	\begin{tabular}{ |p{3cm}|p{9cm}|  }
		\hline
		Company & Count \\
		\hline
		Facebook & Users like \textbf{4,166,667} posts.\\
		\hline
		Twitter & Users send \textbf{347,222} tweets.\\
		\hline
		Youtube & Users upload \textbf{300} hours of video.\\
		\hline
		Instagram & Users like \textbf{1,736,111} photos. \\
		\hline
		Pinterest & Users pin \textbf{9,722} images.\\
		\hline
		Apple & Users download  \textbf{51,000} apps.\\
		\hline
		Netflix & Subscribers stream  \textbf{77.160} hours of video.\\
		\hline
		Reddit & Users cast  \textbf{18,327} votes.\\
		\hline
		Amazon & Receives  \textbf{4,310} unique visitors.\\
		\hline
		Vine & Users play  \textbf{1,041,666} videos.\\
		\hline
		Tinder & Users swipe  \textbf{590,276} times.\\
		\hline
		Snapchat & Users share  \textbf{284,722} snaps.\\
		\hline
		Buzzfeed & Users view  \textbf{34,150} videos.\\
		\hline
		Skype & Users make  \textbf{110,040} calls.\\
		\hline
		Uber & Passengers take  \textbf{694} rides.\\
		\hline

@@

Deep Learning, which itself is a sub-category of Machine Learning, has gained momentum in solving many of the most complicated AI problems. In many fields such as computer vision and voice recognition Deep Learning outperforms best conventional models and techniques both in accuracy and speed. The major distinguishing difference between most of Machine Learning techniques and that of Deep Learning is in feature selection stage. In conventional Machine Learning, labeled/unlabeled data is fed into a set of hand crafted feature detectors. A classifier takes the features as inputs to perform some statistical analysis and decide on the class of each input. Since the engineering of a good set of such hand crafted features is extremely costly and require a lot of expertise and domain knowledge, Deep Learning is introduced as a solution to find and tune such features automatically using machine and data for a particular task.\\

There is a general consensus that, increasing the size of either or both if training data and/or model size will increase the accuracy of prediction in Deep Learning. Therefore, as the size of training data or the model increases, we require to parallelize the process on many machines and processing cores to enable the scaling. Consequently, two parallelization schemes is offered namely, data parallelization and model parallelization.\\

In data parallelization scheme, the training set is cut into smaller data fragments. Each of these fragments is assigned to a different processing node. Each node replicates the global model so that each model will have a fragment of data with a copy of full model. During the training process the model is updated. Since different copies of the model are updated separately using different fragmentations of data, we need to synchronize all copies of the model into a global model that is representative of all the data. The issue with such synchronizations is the large communication overhead. Models are usually relatively huge (currently up to Gigabytes) that we cannot easily synchronize them frequently. In addition, it is also so hard to fit large models in GPUs' memory.\\

\begin{figure}[h]
	\includegraphics[scale=0.33]{./images/data_parallel.png}
	\centering
	\caption{Data parallelism performing Gradient Descent (GD). Model replicas at each node calculate gradients $\Delta w $ on a piece of data. Gradients are later transmitted to global model server to calculate updated weights.}
	\label{fig:data_parallelism}
\end{figure}
In model parallelization scheme, models are split into smaller pieces, which allows for scaling of the model. A drawback of model parallelization is that one has to synchronize all copies of the model even more frequently than data parallelization. Synchronization should take place for each model at every layer, which for nowadays deep networks can be tens of layers. Thus, interconnects play a very important role when it comes to model parallelization. For instance, a typical neural network architecture with 1 billion nodes may require tens of seconds when using a typical Ethernet networking for synchronization. In contrast, HPC known interconnects such a Infiniband and PCI Express would allow for more efficient networking with higher performance.\\
\begin{figure}[h]
	\includegraphics[scale=0.33]{./images/model_parallel.png}
	\centering
	\caption{An example of model parallelism. Each node store a fragment of the overall model and a whole copy of data.}
	\label{fig:model_parallelism}
\end{figure}

There are many crucial issues to be addressed when thinking about integration of Machine Learning and HPC. Questions such as, aside from accelerating Machine Learning with GPUs and FPGAs, is traditional HPC ready to be the drive force behind data savvy Machine Learning applications? Are current top 500 supercomputers capable of powering new AI era of machine learning and Deep Learning? From software point of view, does conventional HPC require extensive modification to deal with mostly correlated data variables in Machine Learning application?
In the remaining of this section we will briefly introduce some of the ongoing researches and projects concentrated empowering Machine Learning with HPC medium.

\subsection*{Project Adam: Building an Efficient and Scalable Deep Learning Training System \cite{chilimbi2014project}}
In 2012 Google announced that they have trained an unsupervised image recognition neural network with 1 billion connections using 10 million training images, which can recognize pictures of cats from Youtube videos. Their training system consisted of a cluster of 1,000 computers with an aggregate of 16,000 cores \cite{le2013building}.\\

Project Adam by Microsoft could be thought as the successor of Google's project. Adam is built using with 30 times less number of machines compared to previous best image classifier. It also is able to train a image classifier using 14 million images of Imagenet \cite{deng2009imagenet} into 22,000 image categories with almost 2 times better accuracy compared to the last best system. Adam is claimed to be 50 times faster than the state of the art systems.\\
